---
title: "Deakin Study Programing Exercise"
author: "Hannah S. Heinrichs"
date: "12/14/2021"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = "100%")
# DO NOT SOURCE ANYTHING ELSE
source("./src/config.R")
save_plot <- FALSE
```


# Inspection of single dataset

```{r include=FALSE, cache=TRUE}
cdata <- load_data(select_id = "C1045", return_data = TRUE)
cstats <- load_stats(select_id = "C1045", return_data = TRUE)

pdata <- load_data(select_id = "P1045", return_data = TRUE)
pstats <- load_stats(select_id = "P1045", return_data = TRUE)
```

The example data is actigraphy data of a dyad with one client and their partner. 
The data was sampled with `r 1` Hz in 60-second epochs. Participants have different measurement coverage. 

```{r include=FALSE}
cinfo <- timeseries_characteristics(cdata, cstats, verbose = TRUE)
```
The client's epochs start at `r min(cdata$date_time)` and end at `r max(cdata$date_time)`. 
The time series spans `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "DAILY"]` days, including `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "ACTIVE"]` ACTIVE periods, `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "REST"]` REST periods and `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "SLEEP"]` SLEEP periods. The dataset contains `r cinfo$complete_tab$n_obs` observations, `r cinfo$complete_tab$n_complete` complete observations with data in all variables, `r cinfo$complete_tab$t_interval_missing` hours are not covered (`r round(cinfo$complete_tab$perc_missing, 2)`%).
`r nrow(cinfo$na_tab)` periods of at least 1 minute have missing observations in one or more variables.
```{r include=FALSE}
pinfo <- timeseries_characteristics(pdata, pstats, verbose = TRUE)
```

The partner's epochs start later, at `r min(pdata$date_time)` and end at `r max(cdata$date_time)`. 
The time series spans `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "DAILY"]` days, including `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "ACTIVE"]` ACTIVE periods, `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "REST"]` REST periods and `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "SLEEP"]` SLEEP periods. The dataset contains `r pinfo$complete_tab$n_obs` observations, `r pinfo$complete_tab$n_complete` complete observations with data in all variables, `r pinfo$complete_tab$t_interval_missing` hours are not covered (`r round(pinfo$complete_tab$perc_missing, 2)`%).
`r nrow(pinfo$na_tab)` periods of at least 1 minute have missing observations in one or more variables.

Investigations of time periods where one or more variables are missing suggest, failure to records stems from taking of the device, e.g. for showering in the morning, forgetting to put the device on again (several hours missing), or technical failures (several minutes missing).
```{r echo=FALSE}
list(client = cinfo$na_tab, partner = pinfo$na_tab) |>
  dplyr::bind_rows(.id = "group_name") |>
  dplyr::mutate(group_name = str_capitalize(group_name)) |>
  # dplyr::arrange(lubridate::date(.data$start_time)) |>
  dplyr::filter(.data$duration > 5) |>
  dplyr::select(-dplyr::all_of("target")) |>
  dplyr::rename(`Group` = group_name, `Start Time` = start_time, `End Time` = end_time, Duration = duration) |>
  kableExtra::kbl(caption = "Table 1") |>
  kableExtra::footnote("Only periods with missing values of > 5 minutes are reported.") |>
  kableExtra::kable_paper(lightable_options = c("hover", "striped"), full_width = T)
```


```{r include=FALSE}
# config to process dyad data
# anker binning at same time point for dyad
anker <- max(cdata$date_time[1], pdata$date_time[1])
```


## Smoothing

```{r BinComparison, include=FALSE}
# construct relative time at anker, here experiment begin
# calculate in 5-minute bins, in 30-minute bins, in 60 minute bins
binned_cdata <- cdata |>
  dplyr::mutate(
    asleep = as.numeric(.data$interval_status == "REST-S"),
    nrel_date_time = (as.numeric(.data$date_time - anker) / 60)
  ) |>
  calc_multiple_bins(
    target_cols = c(TARGET_COLS, "asleep"),
    rel_col = "nrel_date_time",
    bin_sizes = c(5, 10, 20, 30, 60)
  ) |>
  dplyr::bind_rows()

merged_binned_cdata <- dplyr::bind_rows(binned_cdata) |>
  dplyr::filter(.data$date == unique(.data$date)[2])

p_all_bins <- ggplot(data = merged_binned_cdata, aes(x = .data$time, color = as.factor(.data$bin_size))) +
  geom_line(aes(y = .data$white_light)) +
  facet_grid(facets = .data$date ~ .) +
  scale_y_continuous(trans = "log10", labels = scales::comma) +
  scale_x_time(
    breaks = seq(0, 24, 2) * 60 * 60,
    minor_breaks = (0:24) * 60 * 60,
    limits = c(0, 60 * 60 * 24)
  ) +
  scale_color_brewer(palette = "Set1") +
  labs(
    x = "Time [hh:mm:ss]",
    y = "White light [lux]",
    color = "Bin size [min]",
    title = "White light exposure by day",
    subtitle = "Client data, different bin sizes"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 20, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    ),
    legend.position = "top"
  )
```


```{r PlotBinComparison, echo=FALSE, out.width="100%", fig.cap="Figure 1. White light exposure, log scaled.", fig.height=3, fig.width=12, warning=FALSE, cache=TRUE}
p_all_bins
```
Visual inspection of the time course of white light exposure with different time bins suggests that bin of 20 and even more 30 minutes may capture relevant variation best while smoothing over noisy variation. Smaller bins (e.g. 5 or 10 minutes) heavily depend on occasional extreme values, while larger bins (e.g. 60 minutes) flatten regional changes that may still be of interest. Thus, further analysis is based on 30-minute bins.


## Mean light exposure and activity

```{r Bin30, include=FALSE}
# get data with bin size 30 for client and patient
bin_size <- 30
cdata30 <- binned_cdata[binned_cdata$bin_size == bin_size, ]

binned_pdata <- pdata |>
  dplyr::mutate(
    asleep = as.numeric(.data$interval_status == "REST-S"),
    nrel_date_time = (as.numeric(.data$date_time - anker) / 60)
  ) |>
  calc_multiple_bins(
    target_cols = c(TARGET_COLS, "asleep"),
    rel_col = "nrel_date_time",
    bin_sizes = bin_size
  ) |>
  dplyr::bind_rows()
pdata30 <- binned_pdata[binned_pdata$bin_size == bin_size, ]
```

```{r PlotTimeSeries, include=FALSE, cache=TRUE}
# compute plots for time series
cp_color <- plot_colorlight(cdata30, title = "Light exposure of client\n  in 30-minute bins", log = TRUE, save_plot = save_plot)
cp_white <- plot_whitelight(cdata30, title = "White light exposure of client\n  in 30-minute bins", save_plot = save_plot)
cp_activity <- plot_activity(cdata30, title = "Activity of client in 30-minute bins", save_plot = save_plot)
cp_sleep <- plot_sleep(cdata30, title = "% Sleep recording of client\n  in 30-minute bins", save_plot = save_plot)

pp_color <- plot_colorlight(pdata30, title = "Light exposure of partner\n  in 30-minute bins", log = TRUE, save_plot = save_plot)
pp_white <- plot_whitelight(pdata30, title = "White light exposure of partner\n  in 30-minute bins", save_plot = save_plot)
pp_activity <- plot_activity(pdata30, title = "Activity of partner in 30-minute bins", save_plot = save_plot)
pp_sleep <- plot_sleep(pdata30, title = "% Sleep recording of partner\n  in 30-minute bins", save_plot = save_plot)
```

```{r WeekdayColor, echo=FALSE, fig.cap="Figure 2", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_color, pp_color, ncol = 2)
```

```{r WeekdayWhite, echo=FALSE, fig.cap="Figure 3", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_white, pp_white, ncol = 2)
```

```{r WeekdayActitity, echo=FALSE, fig.cap="Figure 4", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_activity, pp_activity, ncol = 2)
```

```{r WeekdaySleep, echo=FALSE, fig.cap="Figure 5", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_sleep, pp_sleep, ncol = 2)
```


The clients has lower fractions of sleep recording during night time hours and generally less sleep recording than the partner, descriptively confirming sleep disturbance in accordance with the diagnosis.
With respect, to light exposure and activity levels, the client and their partner may differ slightly, but not enough to establish presence of a difference by visual inspection of the time series.


## Regularity light exposure and activity

Regularity of light exposure and activity across the time of data collection be examined visually.


```{r include=FALSE, fig.cap="Figure"}
dyad <- list(client = cdata30, partner = pdata30)

```

```{r}
# plots
autoplot_lists <- lapply(names(dyad), \(data_name){
  # autocorrelation and partial autocorrelation for client
  data <- dyad[[data_name]]
  autocor_list <- get_autocor(data, max_lag = Inf)
  var_names <- names(autocor_list)
  auto_plots <- list()
  partial_plots <- list()

  for (i in seq(autocor_list)) {
    var_name <- sub("_", " ", var_names[i])
    current_autocor <- autocor_list[[i]]
    # autocorrelation plot
    auto_plots[[var_name]] <- plot_correlogram(
      cordata = current_autocor,
      data = data,
      kind = "auto",
      subtitle = str_capitalize(sprintf(
        "%s for %s in %.0f-minute bins",
        var_name, data_name, bin_size
      )),
      plot_xlim = c(0, 200),
      save_plot = save_plot
    )
    # partial autocorrelation plot
    partial_plots[[var_name]] <- plot_correlogram(
      cordata = current_autocor,
      data = data,
      kind = "p-auto",
      subtitle = str_capitalize(sprintf(
        "%s for %s in %.0f-minute bins",
        var_name, data_name, bin_size
      )),
      plot_xlim = c(0, 200),
      save_plot = save_plot
    )
  }
  list(auto = auto_plots, partial = partial_plots)
})
names(autoplot_lists) <- names(dyad)
```

```{r echo=FALSE, fig.width=13, fig.height=3, warning=FALSE, cache=TRUE}
# re-organize plots from list
for (target in names(autoplot_lists[[1]][[1]])) {
  out_grobs <- list()
  index <- 1
  for (group in c("client", "partner")) {
    for (def in c("auto", "partial")) {
      out_grobs[[index]] <-
        autoplot_lists[[group]][[def]][[target]]
      index <- index + 1
    }
  }

  gridExtra::grid.arrange(
    grobs = out_grobs,
    ncol = 4
  )
}
```

Visually, autocorrelations for activity do not differ between client and partner. Activity, as the only endogenous variable, clearly shows a regular circadian rhythm. For the other exogenous variables, white light, red light, blue light, and green light, autocorrelations are less pronounced for the client compared to the partner. Only autocorrelations of red light reveal disrupted circadian rhythm for the client but not for the partner.
Quantitatively, the maximum of autocorrelations in a window between 12 and 36 hours are around 24 hours for both, client and partner, and thus do not suggest any deviation from circadian rhythm (see below).
For detailed quantitative evaluation of regularity and periodicity, other statistical means may be employed, such as time frequency analysis.

```{r message=FALSE, warning=FALSE}
# restructure list of autocorrelation tables into one table
all_autocor <- lapply(dyad, \(data){
  autocor_list <- get_autocor(data, max_lag = Inf)
  dplyr::bind_rows(autocor_list, .id = "target")
}) |>
  dplyr::bind_rows(.id = "group_name") |>
  tidyr::pivot_longer(cols = c("acf", "pacf"), names_to = "cor_def", values_to = "cor_val")

# extract lag (after 12h) with highest autocorrelation / partial autocorrelation
# and compute minute of highest autocorrelation (mid of bin)
all_autocor |>
  dplyr::filter(.data$cor_def == "acf") |>
  dplyr::group_by(group_name, target) |>
  # filter 12h < x < 36h after start_time
  dplyr::filter(
    .data$lag > (12 * 60 / bin_size),
    .data$lag < (36 * 60 / bin_size)
  ) |>
  dplyr::summarise(bin_max_val = .data$lag[which.max(.data$cor_val)]) |>
  dplyr::mutate(
    minute_max_val = bin_max_val * bin_size,
    blob = hms::as_hms(minute_max_val * 60)
  ) |>
  dplyr::arrange(target) |>
  dplyr::mutate(
    group_name = str_snake2human(.data$group_name),
    target = str_snake2human(.data$target)
  ) |>
  dplyr::select(dplyr::all_of(c("target", "group_name", "blob"))) |>
  dplyr::rename(`Group name` = group_name, Target = target, `Time of maximum AC` = blob) |>
  kableExtra::kbl(caption = "Table 2") |>
  kableExtra::footnote("AC: Autocorrelation") |>
  kableExtra::kable_paper(lightable_options = c("hover", "striped"), full_width = T)
```
## Correlation of light exposure and activity

To analyze the relationship between light exposure and activity, cross correlations were computed to examine irregularities in their relationship across time.

Strikingly, there are small deflections of the client's cross-correlations between activity and the all distinct wave lengths (red light, green light, blue light). This coincides with spikes of light exposure with small spikes in the clients light exposure that occur between around 2:00:00 and 6:30:00 (cf. Figure 2). Yet, these deflections may coincide with periods of missing values in the morning listed that may bias for the activity estimate for respective bins (cf. Table 1).

```{r include=FALSE}
# cross correlations of activity and lights
crossplot_lists <- lapply(names(dyad), \(data_name){
  data <- dyad[[data_name]]

  # cross correlations
  crosscor_list <- get_crosscor(data, max_lag = Inf)
  target_names <- names(crosscor_list)
  cross_plots <- list()

  # plotting all targets
  for (i in seq(crosscor_list)) {
    target <- target_names[i]
    current_crosscor <- crosscor_list[[i]]
    # cross-correlation plot
    cross_plots[[target]] <- plot_correlogram(
      cordata = current_crosscor,
      data = data,
      kind = "cross",
      subtitle = str_capitalize(sprintf(
        "%s for %s in %.0f-minute bins",
        sub("_", " ", target), data_name, bin_size
      )),
      plot_xlim = c(-200, 200),
      save_plot = save_plot
    )
  }
  return(cross_plots)
})
names(crossplot_lists) <- names(dyad)
```

```{r echo=FALSE, warning=FALSE, fig.width=15, fig.height=3, cache=TRUE}
# re-orgnize and display plots
# iterate over targets
for (target_name in names(crossplot_lists[[1]])) {
  out_grobs <- list()
  index <- 1
  # iterate over client and partner
  for (group in names(crossplot_lists)) {
    out_grobs[[index]] <-
      crossplot_lists[[group]][[target_name]]
    index <- index + 1
  }

  gridExtra::grid.arrange(
    grobs = out_grobs,
    ncol = 2
  )
}
```

# Exploratory analysis on single dyad/ participants

## Goal

Goal is to analyze the sleep status depending on time, light exposure and partner activity. To aggregate data, variables sleep probability be summarized for interval type = SLEEP periods as single observations.

### Possible approaches/ hypotheses

When looking at 30-minute bins, sleep status over time possibly depends on...

- insomnia diagnosis.
- current light exposure / differential reaction to red light exposure
- current partner-activity.

### Additional thoughts

- bins are correlated over time
- sleep status and activity are correlated and have a common cause.
- each day represents a repeated measurement of sleep 


```{r include=FALSE}
# client and partner data in long format
# create column with partner activity
dyad <- dplyr::bind_rows(list(client = cdata30, partner = pdata30), .id = "group_name")
reverse_dyad <- dyad |>
  dplyr::mutate(
    partner_activity = .data$activity,
    group_name = dplyr::recode(.data$group_name, "client" = "partner", "partner" = "client")
  ) |>
  dplyr::select(dplyr::all_of(c("group_name", "date_time", "partner_activity")))
dyad <- dplyr::left_join(dyad, reverse_dyad, by = c("group_name", "date_time"))

# test
# dyad |>
#   dplyr::filter(.data$time == hms::as_hms("17:24:30")) |>
#   dplyr::select(group_name, date_time, activity, partner_activity)
```

```{r AssumptionTests, echo=FALSE}
# select time where subject of prediction is asleep for the complete bin
dyad_sleep <- dyad |>
  dplyr::filter(.data$asleep == 1) |>
  dplyr::select(dplyr::all_of(c("id", "group", "date_time", "date", "time", "sleep_wake", "activity", "partner_activity", LIGHT_COLS, "asleep")))

# partner activity
dens_act <- ggplot(dyad_sleep, aes(x = partner_activity)) +
  scale_x_continuous(trans = "log10", labels = scales::comma) +
  geom_density() +
  labs(
    x = "Activity [counts]",
    y = "Density [fraction]"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    )
  )

# white light
dens_white <- ggplot(dyad_sleep, aes(x = white_light)) +
  scale_x_continuous(trans = "log10", labels = scales::comma) +
  geom_density() +
  labs(
    x = "White Light [lux]",
    y = "Density [fraction]"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    )
  )

# color light
long_light <- dyad_sleep |>
  tidyr::pivot_longer(cols = c("red_light", "green_light", "blue_light"), values_to = "target_value", names_to = "target_name")
dens_color <- ggplot(long_light, aes(x = target_value, color = target_name)) +
  geom_density() +
  scale_color_manual(
    limits = c("red_light", "green_light", "blue_light"),
    values = c("#CC0000", "#009900", "#0066CC"),
    labels = c("red", "green", "blue")
  ) +
  scale_x_continuous(trans = "log10", labels = scales::comma) +
  labs(
    color = "Color",
    x = "Light [\u00B5W / m\u00B2]",
    y = "Density [fraction]"
  ) +
  theme_bw() +
  theme(
    legend.position = c(0.8, 0.8),
    axis.text.x = element_text(
      angle = 45, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    )
  )

grobs_out <- list(dens_color, dens_white, dens_act)
```

```{r, fig.height=4, fig.width=10, warning=FALSE, cache=TRUE}
gridExtra::grid.arrange(grobs = grobs_out, ncol = 3)
```


```{r eval=FALSE, include=FALSE}
# linear model
model <- "sleep_wake ~ time + red_light + green_light + blue_light + partner_activity"
fit <- lm(model, data = dyad_sleep[dyad_sleep$group == "C",])
summary(fit)

# mixed linear model
model <- "sleep_wake ~ time + red_light + green_light + blue_light + partner_activity + (1|day)"
fit <- lme4::lmer(model, data = dyad_sleep[dyad_sleep$group == "C",])
summary(fit)
```


# Exploratory analysis on complete dataset

## Considerations

### Clustering of the data

- observations are clustered within participants
- participants are clustered within dyads
- participants are clustered within clinical status (client vs. partner)

- assuming no autocorrelation component across within-participant observations, i.e.
one sleep periods is assumed not to influence following period

# To Do

- [ ] do not compute separately for client and partner
- [ ] fix y axes
