---
title: "Deakin Study Programing Exercise"
author: "Hannah S. Heinrichs"
date: "12/14/2021"
output: html_document
---

```{r include=FALSE}
source("./src/main.R")
main()
data <- readRDS("./data/preprocessed/CLIENT/C1045_epochs.rds")
aggstats <- readRDS("./data/statistics/CLIENT/C1045_stats.rds")
binned_data <- inspection(data, aggstats)
```

# Analysis on single dyad/ participants

## Goal

Goal is to analyze the sleep status depending on light exposure, activity 
(and partner activity?). To aggregate data, variables sleep probability be 
summarized for interval type = SLEEP periods as single observations.

Possible approaches/ hypotheses:

Sleep status over time in 10-minute bins depends on...
- insomnia diagnosis.
- current light exposure.
- current partner on partner-activity.
- interaction of insomnia diagnosis and partner-activity.

Additional thoughts:
- Sleep status and activity are correlated and have a common cause.



```{r}

# select sleep time
sleep <- data |>
  dplyr::filter(.data$interval_status == "REST-S") |>
  dplyr::select(dplyr::all_of(c("id", "group", "date_time", "date", "time", "interval_status", "sleep_wake", "activity", light_cols)))

# create 30-minute bins
sleep10 <- sleep |>
  dplyr::mutate(nrel_date_time = (as.numeric(.data$date_time - .data$date_time[1]) / 60)) |>
  bin_series_at_anker(
    target_cols,
    rel_col = "nrel_date_time",
    bin_size = 10
  ) |>
  dplyr::mutate(
    group = sleep$id[1],
    group = sleep$group[1],
    sub_id = sleep$sub_id[1]
  )


model <- "sleep_wake ~ time + partner_activity + white_light + red_light + blue_light + green_light"
fit <- lm(model, data = sleep10)
summary(fit)

# lagged regression?
# GLM? Smoothing?

# Analysis ----------------------------------------------------------------

# https://online.stat.psu.edu/stat510/lesson/8/8.2
# detrending necessary? low-pass filter?
# logistic regression of current light and different lags
# compare odds ratio(?) between C and P

model <- "sleep_wake ~ 1"
res <- glm(model, data = data30, family = binomial)
```


# Analysis on complete dataset

## Considerations

### Clustering of the data

- observations are clustered within participants
- participants are clustered within dyads
- participants are clustered within clinical status (client vs. partner)

- assuming no autocorrelation component across within-participant observations, i.e.
one sleep periods is assumed not to influence following period
