---
title: "Deakin Study Programing Exercise"
author: "Hannah S. Heinrichs"
date: "12/14/2021"
output: html_document
---

```{r include=FALSE}
# DO NOT SOURCE ANYTHING ELSE
source("./src/config.R")
```

# Inspection of one example dataset

The example data is actigraphy data of a dyad with one client and their partner. 

```{r echo=FALSE}
# load C1045
c_expl <- load_data(select_id = "C1045", return_data = TRUE)
cdata <- c_expl$epochs
cstats <- c_expl$aggstats

p_expl <- load_data(select_id = "P1045", return_data = TRUE)
pdata <- p_expl$epochs
pstats <- p_expl$aggstats

data <- rbind(cdata, pdata)
aggstats <- rbind(cstats, pstats)
```
The data was sampled with `r 1` Hz for each epoch and consists of `r nrow(cdata)` epochs from the client, and `r nrow(pdata)` epochs from the partner.

```{r}
# 
```


```{r}
# process dyad

anker <- min(cdata$date_time, pdata$date_time)
# join der Datensätze, hinzufügen der column

```


# Analysis on single dyad/ participants

## Goal

Goal is to analyze the sleep status depending on light exposure, activity 
(and partner activity?). To aggregate data, variables sleep probability be 
summarized for interval type = SLEEP periods as single observations.

Possible approaches/ hypotheses:

Sleep status over time in 10-minute bins depends on...
- insomnia diagnosis.
- current light exposure.
- current partner on partner-activity.
- interaction of insomnia diagnosis and partner-activity.

Additional thoughts:
- Sleep status and activity are correlated and have a common cause.



```{r eval=FALSE, include=FALSE}

# select sleep time
sleep <- data |>
  dplyr::filter(.data$interval_status == "REST-S") |>
  dplyr::select(dplyr::all_of(c("id", "group", "date_time", "date", "time", "interval_status", "sleep_wake", "activity", light_cols)))

# TODO density plots
# assumption tests

# create 30-minute bins
sleep10 <- sleep |>
  dplyr::mutate(nrel_date_time = (as.numeric(.data$date_time - .data$date_time[1]) / 60)) |>
  bin_series_at_anker(
    target_cols,
    rel_col = "nrel_date_time",
    bin_size = 10
  ) |>
  dplyr::mutate(
    group = sleep$id[1],
    group = sleep$group[1],
    sub_id = sleep$sub_id[1]
  )

```


```{r eval=FALSE, include=FALSE}
model <- "sleep_wake ~ time + partner_activity + white_light + red_light + blue_light + green_light"
fit <- lm(model, data = sleep10)
summary(fit)

# lagged regression?
# GLM? Smoothing?

# Analysis ----------------------------------------------------------------

# https://online.stat.psu.edu/stat510/lesson/8/8.2
# detrending necessary? low-pass filter?
# logistic regression of current light and different lags
# compare odds ratio(?) between C and P

model <- "sleep_wake ~ 1"
res <- glm(model, data = data30, family = binomial)
```


# Analysis on complete dataset

## Considerations

### Clustering of the data

- observations are clustered within participants
- participants are clustered within dyads
- participants are clustered within clinical status (client vs. partner)

- assuming no autocorrelation component across within-participant observations, i.e.
one sleep periods is assumed not to influence following period
