---
title: "Deakin Study Programing Exercise"
author: "Hannah S. Heinrichs"
date: "12/14/2021"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = "100%")
# DO NOT SOURCE ANYTHING ELSE
source("./src/config.R")
save_plot <- FALSE
```


# Inspection of single dataset

```{r include=FALSE, cache=TRUE}
cdata <- load_data(select_id = "C1045", return_data = TRUE)
cstats <- load_stats(select_id = "C1045", return_data = TRUE)

pdata <- load_data(select_id = "P1045", return_data = TRUE)
pstats <- load_stats(select_id = "P1045", return_data = TRUE)
```

The example data is actigraphy data of a dyad with one client and their partner. 
The data was sampled with `r 1` Hz in 60-second epochs. Participants have different measurement coverage. 

```{r include=FALSE}
cinfo <- timeseries_characteristics(cdata, cstats, verbose = TRUE)
```
The client's epochs start at `r min(cdata$date_time)` and end at `r max(cdata$date_time)`. 
The time series spans `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "DAILY"]` days, including `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "ACTIVE"]` ACTIVE periods, `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "REST"]` REST periods and `r cinfo$status_tab$n[cinfo$status_tab$interval_type == "SLEEP"]` SLEEP periods. The dataset contains `r cinfo$complete_tab$n_obs` observations, `r cinfo$complete_tab$n_complete` complete observations with data in all variables, `r cinfo$complete_tab$t_interval_missing` hours are not covered (`r round(cinfo$complete_tab$perc_missing, 2)`%).
`r nrow(cinfo$na_tab)` periods of at least 1 minute have missing observations in one or more variables.
```{r include=FALSE}
pinfo <- timeseries_characteristics(pdata, pstats, verbose = TRUE)
```

The partner's epochs start later, at `r min(pdata$date_time)` and end at `r max(cdata$date_time)`. 
The time series spans `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "DAILY"]` days, including `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "ACTIVE"]` ACTIVE periods, `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "REST"]` REST periods and `r pinfo$status_tab$n[pinfo$status_tab$interval_type == "SLEEP"]` SLEEP periods. The dataset contains `r pinfo$complete_tab$n_obs` observations, `r pinfo$complete_tab$n_complete` complete observations with data in all variables, `r pinfo$complete_tab$t_interval_missing` hours are not covered (`r round(pinfo$complete_tab$perc_missing, 2)`%).
`r nrow(pinfo$na_tab)` periods of at least 1 minute have missing observations in one or more variables.

Investigations of time periods where one or more variables are missing suggest, failure to records stems from taking of the device, e.g. for showering in the morning, forgetting to put the device on again (several hours missing), or technical failures (several minutes missing).
```{r echo=FALSE}
list(client = cinfo$na_tab, partner = pinfo$na_tab) |>
  dplyr::bind_rows(.id = "group_name") |>
  dplyr::mutate(group_name = str_capitalize(group_name)) |>
  # dplyr::arrange(lubridate::date(.data$start_time)) |>
  dplyr::filter(.data$duration > 5) |>
  dplyr::select(-dplyr::all_of("target")) |>
  dplyr::rename(`Group` = group_name, `Start Time` = start_time, `End Time` = end_time, Duration = duration) |>
  kableExtra::kbl(caption = "Table 1") |>
  kableExtra::footnote("Only periods with missing values of > 5 minutes are reported.") |>
  kableExtra::kable_paper(lightable_options = c("hover", "striped"), full_width = T)
```


```{r include=FALSE}
# config to process dyad data
# anker binning at same time point for dyad
anker <- max(cdata$date_time[1], pdata$date_time[1])
```


## Smoothing

```{r BinComparison, include=FALSE}
# construct relative time at anker, here experiment begin
# calculate in 5-minute bins, in 30-minute bins, in 60 minute bins
binned_cdata <- cdata |>
  dplyr::mutate(
    asleep = as.numeric(.data$interval_status == "REST-S"),
    nrel_date_time = (as.numeric(.data$date_time - anker) / 60)
  ) |>
  calc_multiple_bins(
    target_cols = c(TARGET_COLS, "asleep"),
    rel_col = "nrel_date_time",
    bin_sizes = c(5, 10, 20, 30, 60)
  ) |>
  dplyr::bind_rows()

merged_binned_cdata <- dplyr::bind_rows(binned_cdata) |>
  dplyr::filter(.data$date == unique(.data$date)[2])

p_all_bins <- ggplot(data = merged_binned_cdata, aes(x = .data$time, color = as.factor(.data$bin_size))) +
  geom_line(aes(y = .data$white_light)) +
  facet_grid(facets = .data$date ~ .) +
  scale_y_continuous(trans = "log10", labels = scales::comma) +
  scale_x_time(
    breaks = seq(0, 24, 2) * 60 * 60,
    minor_breaks = (0:24) * 60 * 60,
    limits = c(0, 60 * 60 * 24)
  ) +
  scale_color_brewer(palette = "Set1") +
  labs(
    x = "Time [hh:mm:ss]",
    y = "White light [lux]",
    color = "Bin size [min]",
    title = "White light exposure by day",
    subtitle = "Client data, different bin sizes"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 20, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    ),
    legend.position = "top"
  )
```


```{r PlotBinComparison, echo=FALSE, out.width="100%", fig.cap="Figure 1. White light exposure, log scaled.", fig.height=3, fig.width=12, warning=FALSE, cache=TRUE}
p_all_bins
```
Visual inspection of the time course of white light exposure with different time bins suggests that bin of 20 and even more 30 minutes may capture relevant variation best while smoothing over noisy variation. Smaller bins (e.g. 5 or 10 minutes) heavily depend on occasional extreme values, while larger bins (e.g. 60 minutes) flatten regional changes that may still be of interest. Thus, further analysis is based on 30-minute bins.


## Mean light exposure and activity

```{r Bin30, include=FALSE}
# get data with bin size 30 for client and patient
bin_size <- 30
cdata30 <- binned_cdata[binned_cdata$bin_size == bin_size, ]

binned_pdata <- pdata |>
  dplyr::mutate(
    asleep = as.numeric(.data$interval_status == "REST-S"),
    nrel_date_time = (as.numeric(.data$date_time - anker) / 60)
  ) |>
  calc_multiple_bins(
    target_cols = c(TARGET_COLS, "asleep"),
    rel_col = "nrel_date_time",
    bin_sizes = bin_size
  ) |>
  dplyr::bind_rows()
pdata30 <- binned_pdata[binned_pdata$bin_size == bin_size, ]
```

```{r PlotTimeSeries, include=FALSE, cache=TRUE}
# compute plots for time series
cp_color <- plot_colorlight(cdata30, title = "Light exposure of client\n  in 30-minute bins", log = TRUE, save_plot = save_plot)
cp_white <- plot_whitelight(cdata30, title = "White light exposure of client\n  in 30-minute bins", save_plot = save_plot)
cp_activity <- plot_activity(cdata30, title = "Activity of client in 30-minute bins", save_plot = save_plot)
cp_sleep <- plot_sleep(cdata30, title = "% Sleep recording of client\n  in 30-minute bins", save_plot = save_plot)

pp_color <- plot_colorlight(pdata30, title = "Light exposure of partner\n  in 30-minute bins", log = TRUE, save_plot = save_plot)
pp_white <- plot_whitelight(pdata30, title = "White light exposure of partner\n  in 30-minute bins", save_plot = save_plot)
pp_activity <- plot_activity(pdata30, title = "Activity of partner in 30-minute bins", save_plot = save_plot)
pp_sleep <- plot_sleep(pdata30, title = "% Sleep recording of partner\n  in 30-minute bins", save_plot = save_plot)
```

```{r WeekdayColor, echo=FALSE, fig.cap="Figure 2", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_color, pp_color, ncol = 2)
```

```{r WeekdayWhite, echo=FALSE, fig.cap="Figure 3", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_white, pp_white, ncol = 2)
```

```{r WeekdayActitity, echo=FALSE, fig.cap="Figure 4", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_activity, pp_activity, ncol = 2)
```

```{r WeekdaySleep, echo=FALSE, fig.cap="Figure 5", fig.height=10, fig.width=12, warning=FALSE, out.width="100%", cache=TRUE}
gridExtra::grid.arrange(cp_sleep, pp_sleep, ncol = 2)
```


The clients has lower fractions of sleep recording during night time hours and generally less sleep recording than the partner, descriptively confirming sleep disturbance in accordance with the diagnosis.
With respect, to light exposure and activity levels, the client and their partner may differ slightly, but not enough to establish presence of a difference by visual inspection of the time series.

For better overview, one day is provided as example:

```{r}

```


## Regularity light exposure and activity

Regularity of light exposure and activity across the time of data collection be examined visually.


```{r include=FALSE, fig.cap="Figure"}
dyad <- list(client = cdata30, partner = pdata30)

```

```{r}
# plots
autoplot_lists <- lapply(names(dyad), \(data_name){
  # autocorrelation and partial autocorrelation for client
  data <- dyad[[data_name]]
  autocor_list <- get_autocor(data, max_lag = Inf)
  var_names <- names(autocor_list)
  auto_plots <- list()
  partial_plots <- list()

  for (i in seq(autocor_list)) {
    var_name <- sub("_", " ", var_names[i])
    current_autocor <- autocor_list[[i]]
    # autocorrelation plot
    auto_plots[[var_name]] <- plot_correlogram(
      cordata = current_autocor,
      data = data,
      kind = "auto",
      subtitle = str_capitalize(sprintf(
        "%s for %s in %.0f-minute bins",
        var_name, data_name, bin_size
      )),
      plot_xlim = c(0, 200),
      save_plot = save_plot
    )
    # partial autocorrelation plot
    partial_plots[[var_name]] <- plot_correlogram(
      cordata = current_autocor,
      data = data,
      kind = "p-auto",
      subtitle = str_capitalize(sprintf(
        "%s for %s in %.0f-minute bins",
        var_name, data_name, bin_size
      )),
      plot_xlim = c(0, 200),
      save_plot = save_plot
    )
  }
  list(auto = auto_plots, partial = partial_plots)
})
names(autoplot_lists) <- names(dyad)
```

```{r echo=FALSE, fig.width=13, fig.height=3, warning=FALSE, cache=TRUE}
# re-organize plots from list
for (target in names(autoplot_lists[[1]][[1]])) {
  out_grobs <- list()
  index <- 1
  for (group in c("client", "partner")) {
    for (def in c("auto", "partial")) {
      out_grobs[[index]] <-
        autoplot_lists[[group]][[def]][[target]]
      index <- index + 1
    }
  }

  gridExtra::grid.arrange(
    grobs = out_grobs,
    ncol = 4
  )
}
```

Visually, autocorrelations for activity do not differ between client and partner. Activity, as the only endogenous variable, clearly shows a regular circadian rhythm. For the other exogenous variables, white light, red light, blue light, and green light, autocorrelations are less pronounced for the client compared to the partner. Only autocorrelations of red light reveal disrupted circadian rhythm for the client but not for the partner.
Quantitatively, the maximum of autocorrelations in a window between 12 and 36 hours are around 24 hours for both, client and partner, and thus do not suggest any deviation from circadian rhythm (see below).
For detailed quantitative evaluation of regularity and periodicity, other statistical means may be employed, such as time frequency analysis.

```{r message=FALSE, warning=FALSE}
# restructure list of autocorrelation tables into one table
all_autocor <- lapply(dyad, \(data){
  autocor_list <- get_autocor(data, max_lag = Inf)
  dplyr::bind_rows(autocor_list, .id = "target")
}) |>
  dplyr::bind_rows(.id = "group_name") |>
  tidyr::pivot_longer(cols = c("acf", "pacf"), names_to = "cor_def", values_to = "cor_val")

# extract lag (after 12h) with highest autocorrelation / partial autocorrelation
# and compute minute of highest autocorrelation (mid of bin)
all_autocor |>
  dplyr::filter(.data$cor_def == "acf") |>
  dplyr::group_by(group_name, target) |>
  # filter 12h < x < 36h after start_time
  dplyr::filter(
    .data$lag > (12 * 60 / bin_size),
    .data$lag < (36 * 60 / bin_size)
  ) |>
  dplyr::summarise(bin_max_val = .data$lag[which.max(.data$cor_val)]) |>
  dplyr::mutate(
    minute_max_val = bin_max_val * bin_size,
    blob = hms::as_hms(minute_max_val * 60)
  ) |>
  dplyr::arrange(target) |>
  dplyr::mutate(
    group_name = str_snake2human(.data$group_name),
    target = str_snake2human(.data$target)
  ) |>
  dplyr::select(dplyr::all_of(c("target", "group_name", "blob"))) |>
  dplyr::rename(`Group name` = group_name, Target = target, `Time of maximum AC` = blob) |>
  kableExtra::kbl(caption = "Table 2") |>
  kableExtra::footnote("AC: Autocorrelation") |>
  kableExtra::kable_paper(lightable_options = c("hover", "striped"), full_width = T)
```
## Correlation of light exposure and activity

To analyze the relationship between light exposure and activity, cross correlations were computed to examine irregularities in their relationship across time.

Strikingly, there are small deflections of the client's cross-correlations between activity and the all distinct wave lengths (red light, green light, blue light). This coincides with spikes of light exposure with small spikes in the clients light exposure that occur between around 2:00:00 and 6:30:00 (cf. Figure 2). Yet, these deflections may coincide with periods of missing values in the morning listed that may bias for the activity estimate for respective bins (cf. Table 1).

```{r include=FALSE}
# cross correlations of activity and lights
crossplot_lists <- lapply(names(dyad), \(data_name){
  data <- dyad[[data_name]]

  # cross correlations
  crosscor_list <- get_crosscor(data, max_lag = Inf)
  target_names <- names(crosscor_list)
  cross_plots <- list()

  # plotting all targets
  for (i in seq(crosscor_list)) {
    target <- target_names[i]
    current_crosscor <- crosscor_list[[i]]
    # cross-correlation plot
    cross_plots[[target]] <- plot_correlogram(
      cordata = current_crosscor,
      data = data,
      kind = "cross",
      subtitle = str_capitalize(sprintf(
        "%s for %s in %.0f-minute bins",
        sub("_", " ", target), data_name, bin_size
      )),
      plot_xlim = c(-200, 200),
      save_plot = save_plot
    )
  }
  return(cross_plots)
})
names(crossplot_lists) <- names(dyad)
```

```{r echo=FALSE, warning=FALSE, fig.width=15, fig.height=3, cache=TRUE}
# re-orgnize and display plots
# iterate over targets
for (target_name in names(crossplot_lists[[1]])) {
  out_grobs <- list()
  index <- 1
  # iterate over client and partner
  for (group in names(crossplot_lists)) {
    out_grobs[[index]] <-
      crossplot_lists[[group]][[target_name]]
    index <- index + 1
  }

  gridExtra::grid.arrange(
    grobs = out_grobs,
    ncol = 2
  )
}
```

# Exploratory analysis

## On single dyad

A single dyad consisting of one client with chronic insomnia and their partner is used as example for analysis of the sleep status depending on time, light exposure and partner activity. 

### Possible approaches/ hypotheses

When looking at 30-minute bins, sleep status over time most likely depends on diagnosis of insomnia. First, one may suspect that particularly in the insomnia patient, i.e. the client, light of specific wave lengths may enhance the probability of waking up. 
Second, insomnia patients may easily be woken up be movement by their bed-partners.


```{r include=FALSE}
# client and partner data in long format
# create column with partner activity
dyad <- dplyr::bind_rows(list(client = cdata30, partner = pdata30), .id = "group_name")
reverse_dyad <- dyad |>
  dplyr::mutate(
    partner_activity = .data$activity,
    group_name = dplyr::recode(.data$group_name, "client" = "partner", "partner" = "client")
  ) |>
  dplyr::select(dplyr::all_of(c("group_name", "date_time", "partner_activity")))
dyad <- dplyr::left_join(dyad, reverse_dyad, by = c("group_name", "date_time"))

# test
# dyad |>
#   dplyr::filter(.data$time == hms::as_hms("17:24:30")) |>
#   dplyr::select(group_name, date_time, activity, partner_activity)

# select time where subject of prediction is asleep for the complete bin
# switch day time and night time for calculation only
# extract number day from start of the experiment
# night time is formated as "fraction of day" in daytime
dyad_sleep <- dyad |>
  dplyr::filter(.data$asleep == 1) |>
  dplyr::mutate(
    group = as.factor(.data$group),
    night_time = as.numeric(hms::as_hms(format((.data$date_time + lubridate::hours(12)), format = "%H:%M:%S"))) / (60*60*24), 
    day = as.numeric(as.Date(.data$date_time) - min(as.Date(.data$date_time)))
    ) #|>
  #dplyr::select(dplyr::all_of(c("id", "group", "date_time", "date", "time", "sleep_wake", "activity", "partner_activity", LIGHT_COLS, "asleep")))

#hms::as_hms(dyad_sleep$time - hms::as_hms("12:00:00"))

```

### Variables

The primary outcome measure, probability of waking up is measured by the fraction of wakefulness during the selected minute span, here 30 minutes. Note, that categorization of sleep status (asleep or awake) is based on temporally consistent activity. Sleep status changes when the participant is at least 10 minutes immobile or mobile, thus sleep status and activity measure are correlated and concurrent, and activity is not considered a meaningful predictor.
Predictors of interest are measurement of red light, green light, blue light, and white light. Additionally, movement of bed partners can be quantified by partner-activity during sleep periods of the client.
Distribution of those variables are log-normal and can be transformed to normal distribution. However, there are high rates of missing values in all variables (red light: `r perc_zero(dyad_sleep$red_light)`, green light: `r perc_zero(dyad_sleep$green_light)`, blue light: `r perc_zero(dyad_sleep$blue_light)`, white light: `r perc_zero(dyad_sleep$white_light)`, partner-activity: `r perc_zero(dyad_sleep$partner_activity)`). Therefore, all variables are included in their in their original form.

```{r AssumptionTests, echo=FALSE}
# partner activity
dens_act <- ggplot(dyad_sleep, aes(x = partner_activity)) +
  scale_x_continuous(trans = "log10", labels = scales::comma) +
  geom_density(color=DFLT_COLORS[4]) +
  labs(
    x = "Activity [counts]",
    y = "Density [fraction]"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    )
  )

# white light
dens_white <- ggplot(dyad_sleep, aes(x = white_light)) +
  scale_x_continuous(trans = "log10", labels = scales::comma) +
  geom_density() +
  labs(
    x = "White Light [lux]",
    y = "Density [fraction]"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    )
  )

# color light
long_light <- dyad_sleep |>
  tidyr::pivot_longer(cols = c("red_light", "green_light", "blue_light"), values_to = "target_value", names_to = "target_name")
dens_color <- ggplot(long_light, aes(x = target_value, color = target_name)) +
  geom_density() +
  scale_color_manual(
    limits = c("red_light", "green_light", "blue_light"),
    values = c("#CC0000", "#009900", "#0066CC"),
    labels = c("red", "green", "blue")
  ) +
  scale_x_continuous(trans = "log10", labels = scales::comma) +
  labs(
    color = "Color",
    x = "Light [\u00B5W / m\u00B2]",
    y = "Density [fraction]"
  ) +
  theme_bw() +
  theme(
    legend.position = c(0.8, 0.8),
    axis.text.x = element_text(
      angle = 45, vjust = 0.5, hjust = 1,
      margin = margin(t = -8, r = 0, b = 15, l = 0)
    )
  )

grobs_out <- list(dens_color, dens_white, dens_act)
```

```{r, fig.height=4, fig.width=10, warning=FALSE, cache=FALSE}
gridExtra::grid.arrange(grobs = grobs_out, ncol = 3)
```

### Model setup and estimation

In order to quantify difference between client and patient, it's suggested to use one model.
A model was fitted with data from client and patient entered a grouping vector group, a transformed night time vector, as well as a red, green, blue, and white light values in their original form into the model. An interaction term for the group and any other variable was modeled to capture the differential reaction of client and partner to light and partner activity.

```{r SimpleLinear, echo=FALSE, results='asis'}
# simple linear model
model <- "sleep_wake ~ group*(night_time + red_light + green_light + blue_light + partner_activity)"
fit <- lm(model, data = dyad_sleep)
#summary(fit)
stargazer::stargazer(fit, ci=TRUE, ci.level=0.95, single.row = TRUE, type="html")
```

### Further considerations

Further aspects should be considered, in confirmatory hypothesis testing, e.g. correlation of subsequent bins, and that each day represents a repeated measurement of the same construct, sleep disturbance.
One approach to account for the repeated measurement nature is to model data in a hierarchical fashion, e.g. using linear mixed models. Yet, because of the scarcity of above zero values in predictor variables, there are to few observations per day, which does not allow to use introduce a clustering variable, e.g. days, into the model. Accordingly, estimation procedures yield singularity warnings.


```{r MixedLinear, eval=FALSE, include=FALSE, results='asis'}
# mixed linear model
mdata_sleep <- dyad_sleep |>
  dplyr::mutate(
    #night_time = as.numeric(dyad_sleep$night_time)/(60*60*24),
    sleep_wake = scale(sleep_wake),
    partner_activity = scale(.data$partner_activity),
    red_light = scale(.data$red_light),
    blue_light = scale(.data$blue_light),
    green_light = scale(.data$green_light),
    white_light = scale(.data$white_light)
    )
# simple example model
model <- "sleep_wake ~ group*partner_activity + (1 |day)"
fit <- afex::mixed(model, data = mdata_sleep, method="PB")
summary(fit)
```

Future analyses may incorporate the complete dataset in their analysis. During model setup, one should consider clustering of the data on different levels: observations are clustered within participants, and participants are clustered in dyads with their respective partners and within diagnosis group (client or partner).
Using the full data set, the influence of light exposure and activity of the individual during wake hours on sleep periods and sleep disturbance could be investigated by using the aggregated data on actie and sleep periods provided by the summaries in the raw dataset.
For the analysis, simplifying assumptions could be made, such as that sleep periods do not influence each other.
