---
title: "Analysis"
author: "Hannah S. Heinrichs"
date: "12/14/2021"
output: html_document
---

```{r include=FALSE}

```


## Goal

Goal is to analyze the sleep status depending on light exposure, activity 
(and partner activity?). To aggregate data, variables sleep probability be 
summarized for interval type = SLEEP periods as single observations.

Possible approaches/ hypotheses:

- nowcasting of sleep based on light exposure in current sleep period
- forecasting of sleep based on light exposure during prior activity
- forecasting of sleep based on activity during during day

## Considerations

### Clustering of the data

- observations are clustered within participants
- participants are clustered within dyads
- participants are clustered within clinical status (client vs. partner)

- assuming no autocorrelation component across within-participant observations, i.e.
one sleep periods is assumed not to influence following period

```{r}
# lagged regression?


# GLM? Smoothing?

# Analysis ----------------------------------------------------------------

# https://online.stat.psu.edu/stat510/lesson/8/8.2
# detrending necessary? low-pass filter?
# logistic regression of current light and different lags
# compare odds ratio(?) between C and P

model <- "sleep_wake ~ 1"
res <- glm(model, data = data30, family = binomial)

```
